A Global Genome-wide Scan with Optimal Cutoff Mining for Emerging Biomarkers in Head and Neck Squamous Cell Carcinoma

### 2021-03-22 main.tex
Updates from Overleaf

% PGK1
Phosphoglycerate kinase 1 (PGK1 or PGK-1), a glycolysis enzyme, is responsive in cisplatin-resistant HNSCC cell line (H-1R). 
The resistance is associated with the up-regulated expression of PGK1, CD55, ATP-binding cassette transporter genes (e.g., MDR1, MRP1, and MRP2), and down-regulated expression Caveolin 1\cite{Nakamura2005}.

@@

\subsubsection{TCIA validation by deep learning}
%https://wiki.cancerimagingarchive.net/display/Public/TCGA-HNSC
%https://cancer.digitalslidearchive.org/#!/CDSA/hnsc/TCGA-BA-4076/TCGA-BA-4076-01A-01-BS1?slide=TCGA-BA-4076-01A-01-BS1.4aef481c-f634-4ffc-9fb3-609f59a8caf9.svs
TCGA Pathology Images => The Cancer Imaging Archive (TCIA) by Deep learning (R/Pytorch/Keras framework)
%數位模擬，能讓臨床醫師和研究人員更快速了解細胞成長後的狀況，而不是花費時間等待細胞的培養。

OFA:  TCIA for whole slide deep learning
IL19 is a candidate marker
TCPA modeling: Y (RNA-seq) ← X (slide represents gene expression)
Apply model on TMUH HNSCC test cohort (with survival outcome):
Y (survival) ← X (slide represents gene expression)
(as IL19 validation, to see is it still a biomarker)
%x The Stanford Tissue Microarray Database (TMAD; x http://tma.stanford.edu) 
%https://tma.im/cgi-bin/ontologyBrowser.pl?anchor=C34447; there is few HNSCC slides

%Manning meap book: 不再是黑盒子
%https://www.manning.com/books/trust-in-machine-learning
Trust in Machine Learning *** interpretable or explainable CNN
By saliency mapping: e.x. diabetes retinopathy, model focused on retinal artery
%https://bmcmedgenomics.biomedcentral.com/articles/10.1186/s12920-019-0624-2
%#XAI
Interpretable deep neural network for cancer survival analysis by integrating genomic and clinical data\cite{Hao2019a}
By PyTorch

%https://www.deeplearningbook.org/
image-based phenotypes could be correlated with their underlying genomics data 
TMA by deep learning (CNN) mRNA expression associated with pathological images
*** TCGA TCIA Pathology Images with deep learning: gene expression on pathological section, or digital tissue microarray, digital TMA
%https://www.biorxiv.org/content/10.1101/2020.03.10.985887v1.full
%https://github.com/Serian1992/ImgBio.
%https://www.cancer.gov/research/areas/diagnosis/artificial-intelligence
% NCI
Aiding the Genomics Characterization of Tumors: 
deep learning could predict commonly mutated genes and mRNA expression from the images
*** A deep learning model to predict RNA-Seq expression of tumours from whole slide images\cite{Schmauch2020}
Preprocessing of whole-slide images. The application of deep-learning algo- rithms to histological data is a challenging problem, particularly due to the high dimensionality of the data (up to 100,000 × 100,000 pixels for a single whole-slide image) and the small size of available datasets. We divided the whole-slide images into squares of 112 × 112 $\mu$m (224 × 224 pixels) called “tiles”, and used the Otsu algorithm62 (as implemented in python package skimage) to select only those containing tissue, excluding the white background. We sampled a maximum of 8000 such tiles from each slide. We then extracted 2048 features from those tiles with a 50-layer ResNet63 pretrained on the ImageNet dataset64 (using the Keras implementation), such that a slide could be represented as a 8000 × 2048 matrix.
For the first phase of this work (transcriptome prediction), we accelerated the training of our models through a simple preprocessing step inspired by simple linear iterative clustering (SLIC)65: we used the k-means algorithm (as implemented in python package libkmcuda) to create 100 clusters (supertiles) of tiles on the basis of tile location on the slide, and we averaged the features of the tiles within each cluster. The use of these supertiles reduces the dimensions of a slide to 100 × 2048. The model was first trained on this reduced dataset, with all the TCGA data. Then, for specific organs, fine-tuning was achieved with full-scale data from the organs concerned only.

@@

%With increased investments, NCI’s efforts to realize AI’s potential will lead to more accurate and rapid diagnoses, improved clinical decision-making, and, ultimately, better health outcomes for patients with cancer and those at risk. Updated: August 31, 2020
%

>>>>

% PGK1
Phosphoglycerate kinase 1 (PGK1 or PGK-1), a glycolysis enzyme, is responsive in cisplatin-resistant HNSCC cell line (H-1R). 
The resistance is associated with the up-regulated expression of PGK1, CD55, ATP-binding cassette transporter genes (e.g., MDR1, MRP1, and MRP2), and down-regulated expression Caveolin 1\cite{Nakamura2005}.
PGK1 is responsive in hypoxia status, since it is one of HIF-1\alpha regulated gene revealed in HNSCC\cite{Peng2011}\cite{Shih2017}.

@@

\subsubsection{Transcriptomic Representation by Deep Learning}
%https://wiki.cancerimagingarchive.net/display/Public/TCGA-HNSC
%https://cancer.digitalslidearchive.org/#!/CDSA/hnsc/TCGA-BA-4076/TCGA-BA-4076-01A-01-BS1?slide=TCGA-BA-4076-01A-01-BS1.4aef481c-f634-4ffc-9fb3-609f59a8caf9.svs
%TCGA Pathology Images => The Cancer Imaging Archive (TCIA) by Deep learning (R/Pytorch/Keras framework)

%數位模擬，能讓臨床醫師和研究人員更快速了解細胞成長後的狀況，而不是花費時間等待細胞的培養。
A deep-learning model, named HE2RNA\cite{Schmauch2020}, could use to encode transcriptomic representation from histological whole slide images of TCGA. 
%The transcriptomic representation of TCGA learned by HE2RNA.
%TCIA is for radiomics deep learning
Even with RNA-Seq data obtained from bulk cells isolated from whole slide images, HE2RNA model has also been able to predict the spatial expression of more than 8,000 genes from the H\&E slide of HNSCC alone.
GDC store pathology Aperio SVS image. An SVS file is a digital slide image file created by an Aperio ScanScope slide scanner or others.
CAMK2N1 and IL19 are candidate markers suggested in our current study.

HE2RNA modeling: Y (RNA-seq) ← X (slide represents gene expression)
Apply model on TMUH HNSCC test cohort (with survival outcome):
Y (survival) ← X (slide represents gene expression)
(as IL19 validation, to see is it still a biomarker)
%x The Stanford Tissue Microarray Database (TMAD; x http://tma.stanford.edu) 
%https://tma.im/cgi-bin/ontologyBrowser.pl?anchor=C34447; there is few HNSCC slides

% TMU joiny IRB
We will introduce TMU HNSCC cohort from our joint biobank.

%https://www.deeplearningbook.org/
image-based phenotypes could be correlated with their underlying genomics data 
TMA by deep learning (CNN) mRNA expression associated with pathological images
*** TCGA TCIA Pathology Images with deep learning: gene expression on pathological section, or digital tissue microarray, digital TMA

%The Cancer Imaging Archive (TCIA) by Deep learning (R/Pytorch/Keras framework)
%https://www.biorxiv.org/content/10.1101/2020.03.10.985887v1.full
%https://github.com/Serian1992/ImgBio.
%https://www.cancer.gov/research/areas/diagnosis/artificial-intelligence
% NCI
Aiding the Genomics Characterization of Tumors: 
deep learning could predict commonly mutated genes and mRNA expression from the images
*** A deep learning model to predict RNA-Seq expression of tumours from whole slide images\cite{Schmauch2020}.
A HE2RNA model should be trained to systematically predict RNA-Seq profiles from whole-slide images\cite{Schmauch2020}.
%Preprocessing of whole-slide images. The application of deep-learning algo- rithms to histological data is a challenging problem, particularly due to the high dimensionality of the data (up to 100,000 × 100,000 pixels for a single whole-slide image) and the small size of available datasets. We divided the whole-slide images into squares of 112 × 112 $\mu$m (224 × 224 pixels) called “tiles”, and used the Otsu algorithm62 (as implemented in python package skimage) to select only those containing tissue, excluding the white background. We sampled a maximum of 8000 such tiles from each slide. We then extracted 2048 features from those tiles with a 50-layer ResNet63 pretrained on the ImageNet dataset64 (using the Keras implementation), such that a slide could be represented as a 8000 × 2048 matrix.
%For the first phase of this work (transcriptome prediction), we accelerated the training of our models through a simple preprocessing step inspired by simple linear iterative clustering (SLIC)65: we used the k-means algorithm (as implemented in python package libkmcuda) to create 100 clusters (supertiles) of tiles on the basis of tile location on the slide, and we averaged the features of the tiles within each cluster. The use of these supertiles reduces the dimensions of a slide to 100 × 2048. The model was first trained on this reduced dataset, with all the TCGA data. Then, for specific organs, fine-tuning was achieved with full-scale data from the organs concerned only.

@@

%With increased investments, NCI’s efforts to realize AI’s potential will lead to more accurate and rapid diagnoses, improved clinical decision-making, and, ultimately, better health outcomes for patients with cancer and those at risk. Updated: August 31, 2020
%
HE2RNA was able to spatialize differentially genes specifically expressed by T cells or B cells, even though discriminating between those two types from their morphology alone is notoriously difficult.
HE2RNA could be a useful tool for pathologists and oncologists to consistently quantify immune infiltration or guide treatment decision in the context of immunotherapy\cite{Schmauch2020}.
The practical implications of predicting gene expression level from H\&E slides should be emphasized. For example, it could predict molecular profiling from pathology slides, such as the hormonal status of breast cancer cases.
Recent studies have shown that mRNA and protein expressions may be poorly correlated due to various factors, such as different half-lives and post-transcription machinery.
